Molecular alignment on graphene surface determines tansport properties of graphene/organic semiconductor transistors

### 2020-04-06 manuscript.tex
added Vadym's corrections, fixed doi's in bib file, each sentence in new line

As for the graphene-based electronic devices initial stages of growth are particularly important since in this stage the orientation of molecules on the substrate are determined.
 The geometrical orientation of the electric dipoles and multipoles depends, in turn on the orientation of the molecule on the substrate\cite{geng-2012}.
 The electric field dipole exhibits a $\propto 1/r^{3}$ dependence on the distance from the dipole and is thus important only for the extremely thin overlayes on graphene.
 In addition it is important that the overlayers are structurally coherent i.
e.
 monocrystalline domains are as large as possible in order to reduce disorder-related reduction in the effects of surface dipoles.
 Indeed, as  we show in this work, charge carrier transport in such material systems is crucially dependent on the molecular orientation at the graphene/OS interface.
 For this type of research it is important to select OS that exhibits high charge carrier mobility per se, in order to preserve high mobility on graphene.
   N,N'-bis(n-octyl)-(1,7\&1,6)-dicyanoperylene-3,4:9,10-bisdicarboximide (PDI8-CN2)
 exhibits a remarkable air stability, solution processability and relatively high electron mobility in thin film form\cite{jung-2014,molinari-2009,piliego-2009a}.  The position of the lowest unoccupied molecular orbital (LUMO) in this material is at 7.1 eV\cite{jung-2014}, which makes it suitable for interfacing with graphene, whose Dirac point is at $~4.5$ eV.

@@

Graphene was obtained by mechanical exfoliation of highly oriented pyrolytic graphite.
 Subsequently graphene flakes were transferred onto $p$-doped silicon wafers covered with 285 nm-thick thermally grown oxide layer.
 Onto such surface PDI8-CN2 layers were fabricated by Organic Molecular Beam Epitaxy in a vacuum chamber with the base pressure of $1\times10^{-8}$ mbar.
 PDI8-CN2 powder without any further purification, was used as a source material.
 The temperature of the substrate during organic semiconductor (OS) growth was 45$^\circ$C.
 Samples with three different PDI8-CN2 coverages were examined.
 The coverage variation was achieved by varying the growth time, which comprised 1, 3 and 4 minutes.
 The thickness of graphene flakes and the morphology of PDI8-CN2 layers were examined using an optical microscope and an Atomic Force Microscope (AFM) operating in non-contact mode in air.
 Data analysis was performed with Gwyddion software\cite{necas-2012}.
 The $I-V$ characteristics measurements of GFETs were performed by using a Keithley 6487 Picoammeter/Voltage Supply and a Keithley 2400 SorceMeter.
 Drain current ($I_{DS}$) was measured by varying the gate voltage ($V_G$) in the range from -50 to +50 V at two values of the source-drain bias: 0.
001 and 0.
01 V.
 All measurements were performed in a nitrogen-filled glovebox with H$_2$O and O$_2$ concentration of less than 10 ppm.
  Carrier mobility was calculated from $I-V$ curves using the equation, where $L$ and $W$ are the dimensions of the conductive channel, d is the thickness of the silicon oxide layer.
 Each sample was measured before and after deposition of PDI8-CN2 on it.

@@

At the PDI8-CN2 coverage of 0.7 the position of CNP is shifted slightly (for $\sim$0.5 V)  to more positive voltages and the current at the CNP is slightly increased relative to the clean graphene.
  More importantly, we note a less pronounced change in the slope of the transfer characteristics at negative gate voltages, compared to the one observed on clean graphene (brown curve).
 The observed behavior of the transport properties can be understood in the context of rather complex morphology of PDI8-CN2 layers on graphene, and with is associated interaction of the PDI8-CN2 molecules with graphene.
 At the sub-monolayer coverage of PDI8-CN2 the hole transport continues to proceed through the channel, relatively unhindered, electron transport, however is enhanced relative to clean graphene (cf.
 brown and blue curve in Figure \ref{fig:3}).

@@

In order to explain the observed behavior, we have to consider the dependence of the electronic structure of organic thin layers on the molecular orientation.
  Anisotropic behavior of charge transport properties of organic thin films are known for some time now\cite{sundar-2004}.
 Less common are detailed studies of the dependence of the position of the electron energy levels at different molecular orientation.
 This is particularly true for $n$-type OSs on graphene.
 Zhang et al.
\cite{zhang-2016d} have performed a systematic photoelectron spectroscopy study coupled to density functional theory (DFT) -based calculations of electrostatic potential surfaces of several organic semiconductors including n-type.
 Their findings demonstrate energy shifts of the electronic structures as high as 0.54 eV between the layers with the molecules oriented upright and flat-lying molecules.
  Crucial for such difference is the existence of the dipole (and quadrupole) electric field arising at different moieties of the molecules.
 For example, in the case of pyrrolo[3,4-c]pyrrole-1,4-dione (DPP)  the calculations show a surface electric dipole due to the electron-withdrawing carbonyl and imine groups pointing  away from the surface when the molecules are oriented upright (shorter molecular axis perpendicular to the surface).
 On the other hand flat-lying DPP molecules exhibit a quadrupole moment that points towards the surface.
 Consequently, the energy of HOMO is closer to the vacuum level in the case of flat-lying molecules relative to upright oriented molecules.
Qualitatively, we can apply these arguments also for the case of PDI8-CN2 molecules.
 In the absence of available calculations of the electrostatic potential over the PDI8-CN2 molecule, we can speculate that the imide groups provide an electric dipole that points away from the surface and therefore increases electron binding energy.
 On the other hand DFT calculations of the electrostatic potential of the perylene core by Geng et al.
\cite{geng-2012} show an abundance of negative charge inside the core, implying  that the $\pi$-electron system above the molecule represents an electron-deficient region, resulting in a surface dipole pointing towards the surface.
 Consequently, the electron binding energy in the layer of flat-lying molecules is lower relative to the upright oriented molecules.
  In the case of 4 minutes exposure the energy band alignment in the resulting structure corresponds to a type-II heterostructure with graphene and the intermediate layer of flat-lying molecules with energy band structure shifted closer to the vaccum level than the outer layer.

>>>>

As for the graphene-based electronic devices initial stages of growth are particularly important since in this stage the orientation of molecules on the substrate are determined.
 The geometrical orientation of the electric dipoles and multipoles depends, in turn on the orientation of the molecule on the substrate\cite{geng-2012}.
 The electric field dipole exhibits a $\propto 1/r^{3}$ dependence on the distance from the dipole and is thus important only for the extremely thin overlayes on graphene.
 In addition it is important that the overlayers are structurally coherent i.e.monocrystalline domains are as large as possible in order to reduce disorder-related reduction in the effects of surface dipoles.
 Indeed, as  we show in this work, charge carrier transport in such material systems is crucially dependent on the molecular orientation at the graphene/OS interface.
 For this type of research it is important to select OS that exhibits high charge carrier mobility per se, in order to preserve high mobility on graphene.
   N,N'-bis(n-octyl)-(1,7\&1,6)-dicyanoperylene-3,4:9,10-bisdicarboximide (PDI8-CN2)
 exhibits a remarkable air stability, solution processability and relatively high electron mobility in thin film form\cite{jung-2014,molinari-2009,piliego-2009a}.  The position of the lowest unoccupied molecular orbital (LUMO) in this material is at 7.1 eV\cite{jung-2014}, which makes it suitable for interfacing with graphene, whose Dirac point is at $~4.5$ eV.

@@

Graphene was obtained by mechanical exfoliation of highly oriented pyrolytic graphite.
 Subsequently graphene flakes were transferred onto $p$-doped silicon wafers covered with 285 nm-thick thermally grown oxide layer.
 Onto such surface PDI8-CN2 layers were fabricated by Organic Molecular Beam Epitaxy in a vacuum chamber with the base pressure of $1\times10^{-8}$ mbar.
 PDI8-CN2 powder without any further purification, was used as a source material.
 The temperature of the substrate during organic semiconductor (OS) growth was 45$^\circ$C.
 Samples with three different PDI8-CN2 coverages were examined.
 The coverage variation was achieved by varying the growth time, which comprised 1, 3 and 4 minutes.
 The thickness of graphene flakes and the morphology of PDI8-CN2 layers were examined using an optical microscope and an Atomic Force Microscope (AFM) operating in non-contact mode in air.
 Data analysis was performed with Gwyddion software\cite{necas-2012}.
 The $I-V$ characteristics measurements of GFETs were performed by using a Keithley 6487 Picoammeter/Voltage Supply and a Keithley 2400 SorceMeter.
 Drain current ($I_{DS}$) was measured by varying the gate voltage ($V_G$) in the range from -50 to +50 V at two values of the source-drain bias: 0.001 and 0.01 V.
 All measurements were performed in a nitrogen-filled glovebox with H$_2$O and O$_2$ concentration of less than 10 ppm.
  Carrier mobility was calculated from $I-V$ curves using the equation, where $L$ and $W$ are the dimensions of the conductive channel, d is the thickness of the silicon oxide layer.
 Each sample was measured before and after deposition of PDI8-CN2 on it.

@@

At the PDI8-CN2 coverage of 0.7 the position of CNP is shifted slightly (for $\sim$0.5 V)  to more positive voltages and the current at the CNP is slightly increased relative to the clean graphene.
  More importantly, we note a less pronounced change in the slope of the transfer characteristics at negative gate voltages, compared to the one observed on clean graphene (brown curve).
 The observed behavior of the transport properties can be understood in the context of rather complex morphology of PDI8-CN2 layers on graphene, and with is associated interaction of the PDI8-CN2 molecules with graphene.
 At the sub-monolayer coverage of PDI8-CN2 the hole transport continues to proceed through the channel, relatively unhindered, electron transport, however is enhanced relative to clean graphene (cf.brown and blue curve in Figure \ref{fig:3}).

@@

In order to explain the observed behavior, we have to consider the dependence of the electronic structure of organic thin layers on the molecular orientation.
  Anisotropic behavior of charge transport properties of organic thin films are known for some time now\cite{sundar-2004}.
 Less common are detailed studies of the dependence of the position of the electron energy levels at different molecular orientation.
 This is particularly true for $n$-type OSs on graphene.
 Zhang et al.
\cite{zhang-2016d} have performed a systematic photoelectron spectroscopy study coupled to density functional theory (DFT) -based calculations of electrostatic potential surfaces of several organic semiconductors including n-type.
 Their findings demonstrate energy shifts of the electronic structures as high as 0.54 eV between the layers with the molecules oriented upright and flat-lying molecules.
  Crucial for such difference is the existence of the dipole (and quadrupole) electric field arising at different moieties of the molecules.
 For example, in the case of pyrrolo[3,4-c]pyrrole-1,4-dione (DPP)  the calculations show a surface electric dipole due to the electron-withdrawing carbonyl and imine groups pointing  away from the surface when the molecules are oriented upright (shorter molecular axis perpendicular to the surface).
 On the other hand flat-lying DPP molecules exhibit a quadrupole moment that points towards the surface.
 Consequently, the energy of HOMO is closer to the vacuum level in the case of flat-lying molecules relative to upright oriented molecules.
Qualitatively, we can apply these arguments also for the case of PDI8-CN2 molecules.
 In the absence of available calculations of the electrostatic potential over the PDI8-CN2 molecule, we can speculate that the imide groups provide an electric dipole that points away from the surface and therefore increases electron binding energy.
 On the other hand DFT calculations of the electrostatic potential of the perylene core by Geng et al.
\cite{geng-2012} show an abundance of negative charge inside the core, implying  that the $\pi$-electron system above the molecule represents an electron-deficient region, resulting in a surface dipole pointing towards the surface.
 Consequently, the electron binding energy in the layer of flat-lying molecules is lower relative to the upright oriented molecules.
  In the case of 4 minutes exposure the energy band alignment in the resulting structure corresponds to a type-II heterostructure with graphene and the intermediate layer of flat-lying molecules with energy band structure shifted closer to the vacuum level than the outer layer.
